---
title: "00_folder_set_up"
author: "G. Perkins"
date: "30/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script will generate the base folders required for the PEM project. 
Run this script first, You will need to name the aoi only.

Note when naming AOI avoid spaces, and use capitals to distinguish between names if desired. i.e DateCreek or WilliamsLake


```{r}

AOI <- "OldFort"

#base directory 
AOI_dir <- file.path(".", paste0(AOI, "_AOI"))
raw_dir <- file.path(AOI_dir, "0_raw_inputs")
derived_dir <- file.path(AOI_dir, "1_map_inputs")


# input and data processing directly 

cov_dir <- file.path(AOI_dir, "1_map_inputs", "covariates")
shape_dir <- file.path(AOI_dir, "0_raw_inputs", "base_layers")
shape_raw_dir <- file.path(AOI_dir, "0_raw_inputs", "base_layers", "raw")
dem_dir <- file.path(AOI_dir, "0_raw_inputs", "dem")
lidar_dir <- file.path(AOI_dir, "0_raw_inputs", "dem", "lidar")
trim_dir <- file.path(AOI_dir, "0_raw_inputs", "dem", "trim")

# sample filepaths
out_path <- file.path(AOI_dir, "2_sample_design", "stage1_StudyDesign")
sampling_raw_folder <- file.path(out_path, "input_raster")
clhs_outpath <- file.path(out_path, "clhs_sample_plans")

training_data <- file.path(AOI_dir, "1_map_inputs", "trainingData")
training_data_clean <- file.path(AOI_dir, "1_map_inputs", "trainingData", "clean")

# model building folders
model_dir <- file.path(AOI_dir, "3_maps_analysis")
model_data <- file.path(AOI_dir, "3_maps_analysis", "models")
model_f <- file.path(AOI_dir, "3_maps_analysis", "models", "forest")

model_data_clean <- file.path(AOI_dir, "1_map_inputs", "trainingData", "clean")



# set up folders if not already exist
folder_set_up <- c(AOI_dir, raw_dir,shape_raw_dir, derived_dir, cov_dir, shape_dir, dem_dir,lidar_dir, trim_dir, out_path, sampling_raw_folder, clhs_outpath,training_data, training_data_clean, model_dir, model_data, model_f )

for(fold in folder_set_up){
  
ifelse(!dir.exists(fold), dir.create(fold, recursive = TRUE), FALSE)

}

```


You will need to save the base data in the following folders: 

1) AOI - save to : "NAMEOFAOI_AOI/0_raw_inputs/base_layers/"
2) RAW LIDAR - save to: "NAMEORAOI_AOI/0_raw_inputs/dem/

Note if you have more up to date data for Base lasyers than those available through the BC data catalogue (ie. Bec Layers) save them in the base_layers folder. 

