---
title: "Raster Baselayer Generation "
author:  "Gen Perkins"
script author: "Gen Perkins"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Once the LiDAR processing and vector baselayers have been generated there are some additional outputs including preparing the base dem rasters and generating bgc raster


```{r packages, include=FALSE}
#install.packages("plyr")

library(plyr)
library(sf)
library(dplyr)
library(raster)
library(stars)
library(fasterize)
#install.packages("devtools")
#devtools::install_github("ColinChisholm/pemgeneratr")
#library("pemgeneratr")


```


## Work-flow

This script relies on having a spatial file (AOI.gpkg) defining the study area (area of interest - AOI). Firstly the area of interest is defined along with input and output folders. 

1) generate template raster for multiple scales
2) generate base DEM Layer from 
 - lidar processing (1m)
 - trim (25)
 
3) output dem raster at multiple scales 
4) 



```{r Parameters}
#AOI <- "DateCreek"
#AOI <- "Baboon"
#AOI <- "PeterHope"
#AOI <- "OldFort"
#AOI <- "Buck"
AOI <- "Deception"
#AOI <- "BoundaryTSA"
#AOI <- "Wetzinkwa"
#AOI <- "EagleHills"


AOI_dir <- file.path(paste0(AOI, "_AOI"))
cov_dir <- file.path(AOI_dir, "1_map_inputs", "covariates")
shape_dir <- file.path(AOI_dir, "0_raw_inputs", "base_layers")
dem_dir <- file.path(AOI_dir, "0_raw_inputs", "dem")
trim_dir <- file.path(AOI_dir, "0_raw_inputs", "dem", "trim")
bcdata_dir <- file.path("_data_allBC")
source("./_functions/aoi_snap.R")
source("./_functions/multi_res.R")

```



# 1) create a raster templates for AOI at different scales

```{r}

# read in the aoi
aoi <- st_read(file.path(shape_dir, "aoi_snapped.gpkg")) %>%
  mutate(id = 1)

# create a set of template rasters based on various cell resolution

res <- c(5, 25) # set to 1m and 25m # set to 1m and 25m

for (re in res){
  re = 5

  r_dir <- file.path(cov_dir, paste0(re,"m"))
  
  if (!dir.exists(r_dir)){
  dir.create(r_dir)
    
  } else {
      
    print("Directory already exists!")
  }
  
  if (file.exists(file.path(r_dir, "template.tif"))){
    
    print("template for this resolution already exists!")
  
    } else {
  
  rtemp <- raster(aoi, res = re)
  r_out <- fasterize(aoi, rtemp, field = "id")
  
  writeRaster(r_out, file.path(r_dir, "template.tif"))
    }
}

```


# 2) Generate base DEM Layer from lidar processing to match template rasters. 

```{r}

# generate a 1m DEM from Lidar or 25m from TRIM base DEM 
# note if lidar is unavailable use other metrics TRIM, CDED
  res <- 1
  dem_file <- file.path(dem_dir, "lidar", "1m_DTM_mosaic.tif")
  rtemplate <- raster(file.path(cov_dir, paste0(res,"m"), "template.tif"))
  dem <- raster(dem_file)
  crs(dem) <- crs(rtemplate)
  dem <- crop(dem, rtemplate)
  writeRaster(dem, file.path(cov_dir, paste0(res,"m"), "dem.tif"), overwrite = T)

  
  # Generate base raster for Boundary 
  
#  res <- 2.5
#  dem_file <- file.path(dem_dir, "lidar", "Boundary_250cm_BEM_GQ201907300940_TSA.tif")
#  rtemplate <- raster(file.path(cov_dir, paste0(res,"m"), "template.tif"))
#  dem <- raster(dem_file)
#  crs(dem) <- crs(rtemplate)
#  dem <- crop(dem, rtemplate)
#  writeRaster(dem, file.path(cov_dir, paste0(res,"m"), "dem.tif"))

#in_aoi <- aoi
  

```


# Generate multiple resolution dems

Once the 1m dem is correctly cropped to the aoi, we can generate base dems at various resolutions. For sample planning we use 25m.

```{r}
# read in 1m base raster as needed
dem <- raster(file.path(cov_dir, paste0("1m"), "dem.tif"))
#dem <- raster(file.path(cov_dir, paste0("2.5m"), "dem.tif"))

## Generate alternate coarser grain resolutions of the input
multi_res(dem, output = cov_dir, resolution = c(5)) 

# confirms same extent
l <- list.files(cov_dir, pattern = "*.tif", recursive = TRUE, full.names = TRUE)
#l <- l[613:614]

for(i in l){
c <- raster(i) ;   print(i) ;  print("Resolution") ;  print(res(c)) ; print(as.vector(extent(c)))
  
}


###################################################################################
# tempory fix for the Deception data set: 

# read in 1m base raster as needed
res = 5
dem <- raster(file.path(cov_dir, paste0("5m"), "dem.tif"))
rtemplate <- raster(file.path(cov_dir, paste0(res,"m"), "aspect.tif"))

#folders <- "height_percentiles"
#folders <- "height_metrics"
folders <- "bincentiles"
folders <- "canopy_cover"
  
#ftc <- list.files(file.path(dem_dir, "lidar","lasmetrics_5m","vci"), full.names = T)
ftc <- list.files(file.path(dem_dir, "lidar","lasmetrics_5m",folders), full.names = T)
ftc <- list.files(file.path(dem_dir, "lidar"), full.names = T)


#dem <- raster(file.path(dem_dir, "lidar","lasmetrics_5m","vci", "vc1_mosaic.tif"))
# p90_mosaic, and cov_gap_mosaic

for(i in ftc) {
  i = ftc[4]
  fname = gsub(".tif", "_rproj.tif",basename(i))
  print(fname)
  ff <- raster(i)
  crs(ff) <- crs(rtemplate)

  #stack(ff, rtemplate)
  
  # Use the expanded aoi to create a wider AOI raster
  #extent(rtemplate)

  dem_reproj <- crop(ff, rtemplate)
  dem_aoi_prj <- projectRaster(dem_reproj, rtemplate, method = 'bilinear')

  writeRaster(dem_aoi_prj, file.path(dem_dir, "lidar","lasmetrics_5m", folders, fname), format = "GTiff", overwrite = TRUE)
  
  #stack(dem_aoi_prj, rtemplate)
}

# confirms same extent
l <- list.files(file.path(cov_dir,"5m"), pattern = "*.tif", recursive = TRUE, full.names = TRUE)
l <- l[499:520]

for(i in l){
c <- raster(i) ;   print(i) ;  print("Resolution") ;  print(res(c)) ; print(as.vector(extent(c)))
  
}

stack(l)

##################################################################################

```


## Generate BGC Rasters 

Note in this case the bgc layer on the BCGW is not the latest available so we use an updated version. 
This script causes problems with larger raster sets (5m Boundary will not automatically convert as stars creates a proxy object due to the size of the data)

```{r}
res <- c(5, 10, 25) # set to 1m and 25m
res <- 5

for (re in res) {
  #bec_name <- "dc_becV12_2.gpkg"
  bec_name <- "bec.gpkg"
  
# select the bec shapefile and check format

  bec_sf <- st_read(file.path(shape_dir, bec_name))
  bec_sf <- st_cast(bec_sf, "MULTIPOLYGON") %>%
   mutate(BGC_LABEL = gsub(" ","", BGC_LABEL)) %>%
   mutate(BGC_no = seq(1, length(bec_sf$BGC_LABEL),1))
  
# select resolution of raster 
  rtemplate <- stars::read_stars(file.path(cov_dir, paste0(re,"m"), "template.tif"), proxy = FALSE)

  out <- stars::st_rasterize(bec_sf["BGC_LABEL"],
                           template = rtemplate)
  st_crs(out) = st_crs(rtemplate)

 # mapview::mapview(out)

  stars::write_stars(out,
                   file.path(cov_dir, paste0(re,"m"), "bgc.tif")) #tile name

  
  ## for very large raster (5m boundary dataset we need to use another method)
  
  #rtemplate <- raster::raster(file.path(cov_dir, paste0(re,"m"), "dem.tif"))
  #out <- fasterize(bec_sf, rtemplate, field = "BGC_no")
  #raster::writeRaster(out,
  #                 file.path(cov_dir, paste0(re,"m"), "bgc.tif")) #tile name

 
}
  
```

# 2) Generate base DEM Layer from TRIM/CDED 

Note to generate the landscape layer froim TRIM and expand the area to ensure the entire study area will be included study area 

```{r}
# update to cded function from bcmaps

res = 25
trim_folder <- file.path(cov_dir, "25m_trim")

# create trim folder
ifelse(!dir.exists(trim_folder), dir.create(trim_folder, recursive = TRUE), FALSE)

# Use the expanded aoi to create a wider AOI raster
# read in the aoi
template <- raster(file.path(cov_dir, "25m_trim", "template.tif"))
extent(template)
# expand the template to wider extent to ensure edges are not cut off for landscape calculations
# extend with a number of rows and columns (at each side)

template_ex <- extend(template, c(100,100))
extent(template_ex)

# convert trim to expanded tempate
trim <- raster(file.path(bcdata_dir, "BC_TRIM", "bc_elevation_25m_bcalb.tif"))
trim_aoi <- crop(trim, template_ex)
trim_aoi_prj <- projectRaster(trim_aoi, template_ex, method = 'bilinear')

writeRaster(trim_aoi_prj, file.path(trim_folder, "dem.tif"), format = "GTiff", overwrite = TRUE)

# crop trim output as place holder for DEM when missing lidar 
trim_aoi <- crop(trim, template)
trim_aoi_prj <- projectRaster(trim_aoi, template, method = 'bilinear')

writeRaster(trim_aoi_prj, file.path(cov_dir, paste0(res,"m_trim"), "dem.tif"), format = "GTiff", overwrite = TRUE)


```


```{r}
# export a 5m TRIM Dem 
res = 5
trim_folder <- file.path(cov_dir, "5m_trim")

# create trim folder
ifelse(!dir.exists(trim_folder), dir.create(trim_folder, recursive = TRUE), FALSE)

# Use the expanded aoi to create a wider AOI raster
# read in the aoi
template <- raster(file.path(cov_dir, "5m_trim", "template.tif"))
extent(template)
# expand the template to wider extent to ensure edges are not cut off for landscape calculations
# extend with a number of rows and columns (at each side)

template_ex <- extend(template, c(100,100))
extent(template_ex)

# convert trim to expanded tempate
trim <- raster(file.path(bcdata_dir, "BC_TRIM", "bc_elevation_25m_bcalb.tif"))
trim_aoi <- crop(trim, template_ex)
trim_aoi_prj <- projectRaster(trim_aoi, template_ex, method = 'bilinear')

writeRaster(trim_aoi_prj, file.path(trim_folder, "dem.tif"), format = "GTiff", overwrite = TRUE)

# crop trim output as place holder for DEM when missing lidar 
trim_aoi <- crop(trim, template)
trim_aoi_prj <- projectRaster(trim_aoi, template, method = 'bilinear')

writeRaster(trim_aoi_prj, file.path(cov_dir, paste0(res,"m_trim"), "dem.tif"), format = "GTiff", overwrite = TRUE)

```


```{r}
# Generate base rasters for Boundary (temp solution 
# NOte this is for Boundary only

res = 25
dem_folder <- file.path(cov_dir, "25m")

# read(dem)
template <- raster(file.path(cov_dir, "25m", "template.tif"))


dah <- raster(file.path(cov_dir, "25m", "DAH_25mclass.tif"))
mrvbf <- raster(file.path(cov_dir, "25m","Boundary_25m_MRVBF64Up2Low6.tif"))
dem <- raster(file.path(cov_dir, "25m", "Boundary_25m_DEM2.tif"))
landclass <- raster(file.path(cov_dir, "25m", "LandForm_Seive10_25m.tif"))

#lays <-stack(dah, mrvbf, dem, landclass)


crs(dem)<- "+proj=aea +lat_0=45 +lon_0=-126 +lat_1=50 +lat_2=58.5 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs"

lay_crop <- crop(dem, template)
dem_aoi <- projectRaster(lay_crop, template, method = 'bilinear')

dem_aoi
 anc_layers <- c("dah_3class.tif", "mrvbf_LS_s.tif", "landform_stt.tif", "slope.tif")

writeRaster(dem_aoi, file.path(cov_dir, "25m", "dah_3class.tif"))
writeRaster(dem_aoi, file.path(cov_dir, "25m", "mrvbf_LS_s.tif"))
writeRaster(dem_aoi, file.path(cov_dir, "25m", "landform_st.tif"))
writeRaster(dem_aoi, file.path(cov_dir, "25m", "dem_ls.tif"))


dem_aoi <- crop(dem, template)
dem_aoi<- projectRaster(dem_aoi, template, method = 'bilinear')


```



```{r}

# test the difference in smoothing 

# options: 
plot(x, interpolate = T)
y = disaggregate(x, 5, method = 'bilinear')
y = focal(x , w = matrix(1,5,5), mean)

# both focal

# behrens paper

res = 5
dem <- raster(file.path(cov_dir, paste0(res,"m"), "dem.tif"), overwrite = T)



# 3x3
r3 <- focal(dem, w=matrix(1/9,nrow=3,ncol=3))

# 5x5 mean filter
r5 <- focal(dem, w=matrix(1/25,nrow=5,ncol=5)) 

# 10x10 mean filter
r10 <- focal(dem, w=matrix(1/100,nrow=5,ncol=5)) 


# Gaussian filter
gf <- focalWeight(dem, 2, "Gauss")
rg <- focal(dem, w=gf)


out_dir ="E:/temp/PEM_DATA/BEC_DevExchange_Work/Deception_AOI/results/"

writeRaster(r3, file.path(out_dir, "r3_filter.tif"))
writeRaster(r5, file.path(out_dir, "r5_filter.tif"), overwrite = TRUE)
writeRaster(r10, file.path(out_dir,"r10_filter.tif"), overwrite = TRUE)


seq filter(
increasing filter size()
larger size resolution
)



```



# Reproject the base layers so same projection can be used in stars stack 

```{r}

## writeout option
cov.dir = "D:/PEM_DATA/BEC_DevExchange_Work/Deception_AOI/1_map_inputs/covariates/5m"
#out.dir = "D:/PEM_DATA/BEC_DevExchange_Work/Deception_AOI/1_map_inputs/covariates/5m/reproj"

rasterfiles <- list.files(cov.dir, full.names = TRUE)


baseraster <- raster::raster(rasterfiles[4])


for (iii in 137:150){
      #  print(iii)
        iii = 205
        #sraster = read_stars(rasterfiles[iii])
        #print(sraster)
        updateraster <- raster::raster(rasterfiles[iii])
        rname <- basename(rasterfiles[iii])
        rname
        crs(updateraster) <- crs(baseraster)
        writeRaster(updateraster,(file.path(cov.dir, "reproj", rname)))
} 

```


# check that all rasters are aligned and can stack

```{r}
## writeout option
cov.dir = "D:/PEM_DATA/BEC_DevExchange_Work/EagleHills_AOI/1_map_inputs/covariates/5m"

rasterfiles <- list.files(file.path(cov_dir, "5m"), full.names = TRUE)

baseraster <- raster::raster(rasterfiles[1])

aa <- rasterfiles
rasterfiles <- aa[1:5]

for(ii in rasterfiles){
  rr <- raster(ii)
  print(ii)
  print(extent(rr))
  
  }

[1] "EagleHills_AOI/1_map_inputs/covariates/5m/std.tif" # 55
[1] "EagleHills_AOI/1_map_inputs/covariates/5m/p95.tif" # 46
[1] "EagleHills_AOI/1_map_inputs/covariates/5m/p90.tif" # 45
[1] "EagleHills_AOI/1_map_inputs/covariates/5m/p75.tif" # 44
[1] "EagleHills_AOI/1_map_inputs/covariates/5m/p50.tif" # 43
"EagleHills_AOI/1_map_inputs/covariates/5m/dem.tif" #8
1] "EagleHills_AOI/1_map_inputs/covariates/5m/cov.tif" #6


rno = 8

foi_stack <- raster(rasterfiles[rno])
rname = rasterfiles[rno]
  rname
#foi_stack <- stack(foi)
#foi_stack

fst <- raster::crop(foi_stack, baseraster)
crs(fst) <- crs(baseraster)

stack(fst, baseraster)

#fst <- resample(fst, baseraster, method = "bilinear")

writeRaster(fst, file.path(rname), bylayer = TRUE, suffix = names(fst), overwrite = TRUE)

 

stack(rasterfiles)



for (iii in 137:150){
      #  print(iii)
        iii = 55
        #sraster = read_stars(rasterfiles[iii])
        #print(sraster)
        updateraster <- raster::raster(rasterfiles[iii])
        rname <- basename(rasterfiles[iii])
        rname
        crs(updateraster) <- crs(baseraster)
        writeRaster(updateraster,(file.path(cov.dir, "reproj", rname)))

        



########################################################

rasterfiles <- list.files(cov.dir, full.names = TRUE)

for (iii in 1:length(rasterfiles)){
        
        iii = 1
        
        orig_raster <- rasterfiles[iii] # 5
        print (orig_raster)
        orig_raster <- raster(orig_raster)
        
        outfile <- gsub("sdat", "tif", sagafile)
        r <- readGDAL(file.path(saga_files,"sagaTmp",sagafile))
        w <- file.path(saga_files, "outputs", outfile) #, sep = "")
        writeGDAL(r, w)
        
        rfile <- raster(w)
        
        if(is.na(crs(rfile))) {
                print("setting projection to 3005")
                # ensure raster is in BC albers projection
                PROJ <- crs(paste("+init=epsg:",3005, sep = "")) 
                crs(rfile) <- PROJ 
                
        } else {
                print("projection already set to 3005")
        }
        
        w_out <- writeRaster(rfile, file.path(saga_files, basename(w)), overwrite = TRUE, driver = "GTiff")
        
}

stack(rasterfiles[1:5])


```

